---
title: "Using the ReIns package"
author: "Tom Reynkens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction


The ReIns package contains functions and datasets of the book "Reinsurance: Actuarial and Statistical Aspects" (2015) of Albrecher, Beirlant and Teugels.

It contains implementations of

* Basic extreme value theory (EVT) estimators and graphical methods as described in "Statistics of Extremes: Theory and Applications" (2004) of Beirlant, Goegebeur, Segers and Teugels.

* EVT estimators and graphical methods adapted for censored and/or truncated data.

* Splicing of mixed Erlang distributions with EVT distributions (Pareto, GPD).

* VaR, expected shortfall and excess-loss premium estimates.

This vignette describes how to use the most important functions of the package.
We split this in several sections: *datasets, graphical methods, estimators for the extreme value index,
estimators for quantiles and return periods, splicing* and *risk measures*.


# Datasets

Three datasets are available: Norwegian fire insurance data (`norwegianfire`), SOA group medical insurance data (`soa`) and Secura RE automobile reinsurance data (`secura`). These datasets were already discussed in Beirlant et al. (2004).

The illustrations will be done using the Norwegian Fire insurance dataset which contains fire insurance claims for a Norwegian insurance company for the period 1972 to 1992. The sizes of the fire insurance claims are expressed in 1000 NOK.

```{r, fig.width=5, fig.height=3.5, fig.align='center'}
library("ReIns")

data("norwegianfire")

# Claim year
year <- norwegianfire$year+1900

# Claim size
size <- norwegianfire$size

# Plot norwegian fire insurance data per year
plot(year, size, xlab="Year", ylab="Size")
```

# Graphical methods

QQ-plots and their derivative plots are an essential part of extreme value theory. We focus on three important types: the Exponential QQ-plot, Pareto QQ-plot and log-normal QQ-plot.

## Exponential QQ-plot

The exponential QQ-plot can be easily drawn using `ExpQQ`. Its derivative plot, also dubbed mean excess plot is key in determining what type of distribution the data comes from. The mean excess values $e_{k,n}$ are plotted using `MeanExcess` and one has the choice to plot them versus the order statistics $X_{n-k,n}$ (`k=FALSE`) or versus the number of exceedances $k$ (`k=TRUE`).

In this case we see that the mean-excess plot is more or less linearly increasing as a function of  $X_{n-k,n}$ which indicates that the data may come from a Pareto distribution. The exponential QQ-plot is not linear at all but concave which reinforces the conclusion drawn from the mean excess plot.

```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Exponential QQ-plot
ExpQQ(size)


# Mean excess plot with X_{n-k,n}
MeanExcess(size)

# Mean excess plot with k
MeanExcess(size, k=TRUE)
```

## Pareto QQ-plot
Taking the logarithm of Pareto distributed data gives exponentially distributed data, so the Pareto QQ-plot and
the exponential QQ-plot are closely related. Using the commands `ParetoQQ` and `ParetoQQ_der`, the Pareto QQ-plot and its derivative plot can be drawn. Note that these derivatives are nothing more than the Hill estimates.

The Pareto QQ-plot is now linear which indicates that the Pareto distribution is suitable. The derivative plots can be used to estimate the tail index $\gamma=1/\alpha$ of the Pareto distribution (cf. Hill estimator).
```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Pareto QQ-plot
ParetoQQ(size)


# Derivative plot with log X_{n-k,n}
ParetoQQ_der(size)

# Derivative plot with k
ParetoQQ_der(size, k=TRUE)
```


## Log-normal QQ-plot
Finally, one can also consider the log-normal QQ-plot (`LognormalQQ`) and its derivative plot (`LognormalQQ_der`).
It is clear that a log-normal distribution is not suitable for this data.


```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Log-normal QQ-plot
LognormalQQ(size)


# Derivative plot with log X_{n-k,n}
LognormalQQ_der(size)

# Derivative plot with k
LognormalQQ_der(size, k=TRUE)
```

# Estimators for the EVI

## Estimators for $\gamma>0$
The most famous estimator for the EVI $\gamma$ is the Hill estimator which can be obtained by fitting the Pareto distribution to the relative excesses $x/x_{n-k,n}$ using Maximum Likelihood Estimation (MLE). The typical Hill plot can be made using `Hill`.

The bias of the Hill estimator can be problematic so one can consider a bias-reduced estimator which uses a regression type approach `Hill.2oQV`. An other solution is to use the EPD estimator (`EPD`) which fits the extended Pareto distribution instead of the ordinary Pareto distribution to the relative excesses.

Using these bias-reduced estimators can give an idea about a good choice for $k$. Suitable choices of $k$ are values where the two plots intersect and which are not too low (otherwise the variance is too high). This yields an estimate for $\gamma$ around 0.75 (and $k$ around 3000).


```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Hill plot
H <- Hill(size, plot=TRUE, col="blue")

# Add EPD estimator
EPD(size, add=TRUE, col="orange", lty=2)

legend("bottomright",c("Hill","EPD"), col=c("blue","orange"), lty=1:2)
```

## General estimators
The previous estimators can only be used when $\gamma$ is strictly positive. Therefore, the generalised QQ-plot was proposed. This QQ-plot is a generalisation of the Pareto QQ-plot and can also have negative slopes. The function `genQQ` needs the Hill estimates as input.

We see that the generalised QQ-plot is strictly increasing which indicates a strictly positive $\gamma$.

```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Generalised QQ-plot
genQQ(size, gamma=H$gamma)
```


A generalised Hill estimator can then be defined which is an estimator for the slope of the generalised QQ-plot.
Alternatively, one can also consider the moment estimator `Moment` and the Peaks-Over-Threshold estimator (`GPDmle`) which fits the Generalised Pareto Distribution (GPD) to the excesses $x-x_{n-k,n}$ using Maximum Likelihood Estimation (MLE). The POT can however be very time consuming on large datasets!

The generalised QQ-plot and the moment estimator indicate that values for $\gamma$ around 0.75 are indeed suitable.

```{r, fig.width=5, fig.height=3.5, fig.align='center'}

# Generalised Hill estimator
genHill(size, gamma=H$gamma, plot=TRUE, col="blue", ylim=c(0.5,0.8))

# Moment estimator
Moment(size, add=TRUE, lty=2)

legend("bottomright",c("genHill","Moment"), col=c("blue","black"), lty=1:2)
```



